---
title: "NYC Squirrel Data"
author: "Rao"
date: "2023-07-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading data from files
The `read` family functions are useful for reading text-based data formats like csv (comma-separated value), tsv (tab-separated value) and txt.

```{r}
# reading data from file
squirrel <- read.csv("nyc_squirrel_uniq_plus.csv")

```

The data that is read is stored in the form of a `data.frame`, which is a way to represent tabular data in R.

## A quick look at the data.frame
Several R functions enable us to peek at the data, as well as get information about the data (e.g. the number of data points, the kind of data)

```{r}
head(squirrel)
tail(squirrel)
View(squirrel)


dim(squirrel)
nrow(squirrel)
ncol(squirrel)
colnames(squirrel)
rownames(squirrel)
```

## Subsetting a data.frame

```{r}
squirrel[1:5, ]
squirrel[ , 1:5]
squirrel[1:5, 1:5]
squirrel$Age
squirrel[ , "Unique.Squirrel.ID"]

```


## Filtering

Keeping only tha part of the data that satisfy a specific condition (or a combination of conditions) is called filtering. Filtering in base R can be done using the subsetting notation `[ ]`.

```{r}
squirrel[squirrel$Age == "Adult", ] # only adult squirrels
squirrel[squirrel$Shift == "AM", ] # only found in the morning

# Filtering for only adult squirrels found in the morning
squirrel[(squirrel$Age == "Adult") & (squirrel$Shift == "AM"), ]
```

## Summarising a data.frame

The summary function is very helpful to find out the summary statistics of a data.frame column-by-column.

```{r}
summary(squirrel)

table(squirrel$Shift)
```

## Exercise

- How many gray squirrels were found in the morning (AM) vs afternoon (PM)?

- Other questions from Etherpad


