---
title: "NYC Squirrel Data - Tidyverse"
author: "Rao"
date: "2023-07-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Installing packages

To install packages from CRAN:

```{r}
install.packages("tidyverse")
library(tidyverse)
```
## Reading data from files
The `read` family functions are useful for reading text-based data formats like csv (comma-separated value), tsv (tab-separated value) and txt.

```{r}
# reading data from file
squirrel <- read.csv("../datasets/nyc_squirrel_uniq_plus.csv")

```

The data that is read is stored in the form of a `data.frame`, which is a way to represent tabular data in R.


## Subsetting a data.frame

```{r}
squirrel %>%
  head()

squirrel %>%
  select(Unique.Squirrel.ID)

```


## Filtering

Keeping only tha part of the data that satisfy a specific condition (or a combination of conditions) is called filtering. Filtering in base R can be done using the subsetting notation `[ ]`.

```{r}
squirrel %>%
  filter(Age == "Adult")
```

## Summarising a data.frame

The summary function is very helpful to find out the summary statistics of a data.frame column-by-column.

```{r}
squirrel %>%
  group_by(Age) %>%
  summarise(mean(Weight))

```

## Exercise

- How many gray squirrels were found in the morning (AM) vs afternoon (PM)?

- Other questions from Etherpad


