
---
title: "Getting started in R"
author: "Rao"
date: "2023-07-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Printing text or numbers

R is a statistical programming language. At the most basic level, you can use it to print text and numbers.

```{r}
"apple"
"a longer piece of text"
1
```
R can be used as a calculator.
```{r}
1 + 4
9 - 10
53 * 2
4/2
3^2
5 %% 2
```
## Commenting and uncommenting
Comments in code help to document what you are doing with the data.
```{r}
# addition (block comment)
1 + 4

9 - 10 # subtraction (inline comment)
```

## Using functions in R
We can use the print function to do the same thing. The print function is being used in the above code blocks too, but behind the scenes. Below, we use it more explicitly.

```{r}
print("apple")
print("a longer piece of text")
print(1)
print(10^4)
```

A function is usually a word followed by parentheses. Within the parentheses, you can provide 0 or more arguments. Arguments either provide the function with some essential information that is needed for it to run, or modify the way the function works.

```{r}
help()
citation()
```

## Variables

We can give names to pieces of data. Data is assigned to a variable using the <- or = symbols. They are interchangeable, but it's a good idea to be consistent.

```{r}
new_variable <- 1234
myData <- 10 - 2
numero_uno <- 45
numero_dos <- 55
print(numero_uno + numero_dos)
numero_tres <- numero_uno + numero_dos # this doesn't print anything
print(numero_tres)
```


## Exercise
```{r}
# Find out what additional arguments are available in the print() function
# You can use the help() function, ? or Google search

```